Magnetic Inhibition of Convection and the Fundamental Properties of Low--Mass Stars

### 2016-02-28 paper/ms.tex
small edits to magnetic turn off section

{\bf Observational errors may also provide a plausible explanation for the shallower slope of the HR diagram in the vicinity of 5\,000~K. Below this temperature, stellar atmospheric opacity becomes dominated by molecular species, particularly TiO in the optical and H$_2$O in the infrared. It can be argued that the effective temperature scale used to determine the properties of K- and M-dwarfs is too cool by about 200 K. \citet{Preibisch1999} constructed an spectral type to effective temperature transformation at a gravity intermediate between dwarf and giant scales using the transformation derived for luminosity class IV objects by \citet{deJager1987}. Comparing their results to the 5 -- 30 Myr pre-main-sequence star temperature scale proposed by \citet{Pecaut2013}, one finds that the temperature scale adopted by \citeauthor{Preibisch1999} is about 100~K cooler for early K-type stars. However, this trend reverses around spectral type M0, with \citeauthor{Preibisch1999} finding temperatures about 100~K warmer than \citet{Pecaut2013} for early-to-late M-type stars. Considering these differences, the sharp knee feature in the HR diagram may be accentuated by the adoption of an erroneous temperature scale, but the general trend that late-K to M-type stars are significantly cooler than theoretical predictions at 10 Myr remains. Furthermore, \citet{Herczeg2015} independently find a temperature scale consistent with \citeauthor{Preibisch1999}'s using PHOENIX BT-Settl models \citep{Allard2011} to infer effective temperatures from infrared photometric colors. The evidence suggests the adopted temperature scale is not the dominant reason for the observed transition in the HR diagram.}

Instead of being the result of observational errors, the shallower slope in the HR diagram may indicate {\bf a transition} where the influence of magnetic inhibition of convection becomes {\bf a critical component governing a star's structure}.

>>>>

{\bf Errors transforming from photometric colors to effective temperatures may also provide a plausible explanation for the shallower slope of the HR diagram in the vicinity of 5\,000~K. Below this temperature, stellar atmospheric opacity becomes increasingly dominated by molecular species, particularly TiO in the optical and H$_2$O in the infrared, which poses problems when attempting to derive a robust effective temperature scale. If the effective temperature scale adopted by \citet{Preibisch1999} is too cool by about 200 K, the observed transition might disappear and low-mass K- and M-type stars would fall closer to the 10 Myr standard model isochrone.

\citet{Preibisch1999} constructed an spectral type to effective temperature transformation at a gravity intermediate between dwarf and giant scales using the transformation derived for luminosity class IV objects by \citet{deJager1987}. This temperature scale is consistent with the more recent scale of \citet{Herczeg2015} who used PHOENIX BT-Settl model atmospheres \citep{Allard2011} to infer effective temperatures from photometric colors. Agreement between empirical methods and theoretical predictions is reassuring. However,
comparing these results to the 5 -- 30 Myr pre-main-sequence star temperature scale proposed by \citet{Pecaut2013}, one finds that the temperature scale adopted by \citeauthor{Preibisch1999} is about 100~K cooler for early K-type stars, but the two temperature scales agree around spectral type M0. For early-to-late M-type stars, \citeauthor{Preibisch1999} find temperatures about 100~K warmer than \citet{Pecaut2013}, a reversal from the comparison for late-K-type stars. Considering these differences, the sharp knee feature in the HR diagram may be accentuated by the adoption of an erroneous temperature scale, but it does not appear sufficient to fully explain the shallower slope observed in the HR diagram.}

Instead of being the result of errors in the data, the shallower slope in the HR diagram may indicate {\bf a real transition} where the influence of magnetic inhibition of convection becomes {\bf a more critical component governing a star's structure}.
